Stochastic Variance Reduced Policy Gradient

### 2018-01-24 svrpg.tex
added some explanations to proof
changed minor details in proofs; next:minibatch

\begin{assumption}\label{ass:M2}
	There is a constant $M_2<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
\[
	\mathbb{V}ar\left(\omega(\tau)\right) \leq (M_2-1)
\]
\end{assumption}

@@

\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
Under Assumptions \ref{ass:bounded_score} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some constants $\psi>0, D>0$ and $F>0$ and for proper choice of step sizes $\alpha_t$, the following property:
\begin{align*}
	&\EVV[]
	{\norm[]{\nabla J(\vtheta_A)}^2} 
		\leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} \\
		&\qquad+
		\frac{D\max_s\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}}{\psi N}
		+\frac{F(M_2-1)}{\psi}
\end{align*}
\end{restatable}

@@

%Lemma 2
\begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
Under Assumption \ref{ass:bounded_score}
, the expected squared norm of the SVRPG gradient can be bounded as follows:
\[
\Ets{\gradBlack{\vtheta_t^{s+1}}} \leq
\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
\nonumber 
+dG^2R^2\left(M_2-1\right)
\]
\end{restatable}
\begin{proof}

	\begin{align}
	\Ets{\gradBlack{\vtheta_t^{s+1}}} 
	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
			+ \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
			+ \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)-\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
			+\gradJ{\vtheta_t^{s+1}} - \gradJ{\tilde{\vtheta}^s}}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
	+\Ets[]{\norm[]{\gradApp{\tilde{\vtheta}^s}{N} - \Ets[]{\gradApp{\tilde{\vtheta}^s}{N}}}^2} \nonumber\\
	&\qquad+ 
	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)- \Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\\
	&\qquad+ 
	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)- \Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\
	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
			+\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber
	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2}\Ets{(\omega(\tau_i)-1)^2} \\
	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2}\left(\Ets{\omega^2(\tau_i)}-1\right) \nonumber\\
	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2}\left(M_2-1\right),
\end{align}
where (1) is from the definition of $\gradApp{\vtheta}{N}$, (2) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (3) is from L-smoothness, 
(4) is from Assumption \ref{ass:bounded_score}, and (5) is from Assumption \ref{ass:M2}.
\end{proof}

%Lemma 0
\begin{restatable}[]{lemma}{auxzero}\label{lemma:aux0}
Under Assumption \ref{ass:bounded_score},
the expected scalar product between $\gradBlack{\vtheta_t^{s+1}}$ and any function $f(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$ can be bounded as follows:
\begin{align*}
\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} &\geq
\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \\
&\qquad-
\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2}
\end{align*}
\end{restatable}
\begin{proof}
\begin{align}
	\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
	&=
	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\dotprod{\Ets{e_s}}{\Ets{f(\vtheta_t^{s+1})}}
	+\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}  \nonumber\\
	&= 
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})} \\
	&\geq  
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad-
	\Tr\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right| \nonumber \\
	%
	&\geq
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad-\Tr\left(\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{f(\vtheta_t^{s+1})}}\right) \nonumber\\
	%
	&\geq	
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad-
	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} -\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\\
	%
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad-
	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \\
	%
	&\geq
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad-
	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2},\nonumber
\end{align}
where (6) is from Definition \ref{def:ideal}; (7) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (8) is from $\Ets{e_s}=0$; (9) is from Young's inequality; and (10) is from the definition of $\gradApp{\vtheta}{N}$.
\end{proof}

%Lemma 1
\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
Under Assumptions \ref{ass:bounded_score} ans \ref{ass:M2}, the expected squared norm of the true gradient $\gradJ{\vtheta_t^{s+1}}$, for appropriate choices of $\alpha_t\geq0$ and $\beta_t>0$, can be bounded as follows:
\[
	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_t\Tr\Varts{\gradApp{\tilde{\vtheta}^s}{1}}}{N\Psi_t}
	+\frac{f_t(M_2-1)}{\Psi_t},
\]
	where
\begin{align*}
	&R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \\
	&c_{m-1} = 0, \\
	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\alpha_t^2\right) + \alpha_tL^2+\frac{\alpha_t^2L^3}{2}, \\
	&\Psi_t = \alpha_t\left(1-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right), \\
	&f_t = \alpha_t^2\frac{G^2R^2(L+2c_{t+1})}{2(1-\gamma)^2}.
\end{align*}
More precisely, the following constraints on $\alpha_t$ and $\beta_t$ are sufficient:
\begin{align*}
&0\leq\alpha_t < \frac{2(1-\nicefrac{c_{t+1}}{\beta_t})}{L+2c_{t+1}} \\
&\beta_t > c_{t+1}.
\end{align*}
\end{restatable}
\begin{proof}
	We have:
	\begin{align}
	\Ets{J(\vtheta_{t+1}^{s+1})} 
	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \\
	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
	&\geq
	\Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
	&\qquad-
	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2},
\end{align}
where (11) is from the L-smoothness of $J(\vtheta)$, (12) is from the SVRPG update, and (13) is from Lemma \ref{lemma:aux0}.

@@

	+G^2\frac{R^2}{(1-\gamma)^2}\left(M_2-1\right)\right)

@@

	&\qquad-\alpha_t^2\frac{G^2R^2(L+2c_{t+1})}{2(1-\gamma)^2}(M_2-1) \\

@@

	-f_t(M_2-1),

@@

To complete the proof, besides rearranging terms, we have to ensure that $\Psi_t>0$ for each $t$. This gives the constraints on $\alpha_t$ and $\beta_t$.
\end{proof}

>>>>

\begin{assumption}\label{ass:M2}
	There is a constant $\Omega<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
\[
	\mathbb{V}ar\left(\omega(\tau)\right) \leq \Omega
\]
\end{assumption}

@@

\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
Under Assumptions \ref{ass:bounded_score} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some constants $\psi>0, D>0$ and $F>0$ and for proper choice of step sizes $\alpha_t$, the following property:
\begin{align*}
	&\EVV[]
	{\norm[]{\nabla J(\vtheta_A)}^2} 
		\leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} \\
		&\qquad+
		\frac{D\max_s\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}}{\psi N}
		+\frac{F\Omega}{\psi}
\end{align*}
\end{restatable}

@@

%Lemma 2
\begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
Under Assumption \ref{ass:bounded_score}
, the expected squared norm of the SVRPG gradient can be bounded as follows:
\[
\Ets{\gradBlack{\vtheta_t^{s+1}}} \leq
\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
\nonumber 
+dG^2R^2\left(M_2-1\right)
\]
\end{restatable}
\begin{proof}
	\Ets{\gradBlack{\vtheta_t^{s+1}}} 
	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
			+ \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
	&= \mathbb{E}_{t\mid s}\left[\left\|\gradApp{\tilde{\vtheta}^s}{N}
			+ \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
			\right.\right.\nonumber\\&\qquad\left.\left.
			-\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
			+\gradJ{\vtheta_t^{s+1}} - \gradJ{\tilde{\vtheta}^s}\right\|^2\right] \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
	+\Ets[]{\norm[]{\gradApp{\tilde{\vtheta}^s}{N} - \Ets[]{\gradApp{\tilde{\vtheta}^s}{N}}}^2} \nonumber\\
	\mathbb{E}_{t\mid s}\left[\left\|\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
		\right.\right.\nonumber\\&\qquad\left.\left.
		- \Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}\right\|^2\right] 
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	\\
		\mathbb{E}_{t\mid s}\left[\left\|\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
			\right.\right.\nonumber\\&\qquad\left.\left.
			-\Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}\right\|^2\right]  \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\
	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
			+\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2}\Ets{(\omega(\tau_i)-1)^2} \\
	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2}\left(\Ets{\omega^2(\tau_i)}-1\right) \nonumber\\
	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2\Omega}{(1-\gamma)^2},
where (1) is from the definition of $\gradApp{\vtheta}{N}$, (2) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (3) is from L-smoothness, 
(4) is from Assumption \ref{ass:bounded_score}, and (5) is from Assumption \ref{ass:M2}.
\end{proof}

%Lemma 0
\begin{restatable}[]{lemma}{auxzero}\label{lemma:aux0}
Under Assumption \ref{ass:bounded_score}, for any function $f(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$:
\begin{align*}
\left|\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
-\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
\right|
\leq
\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2}
\end{align*}
\end{restatable}
\begin{proof}
\begin{align}
	\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
	&=
	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\dotprod{\Ets{e_s}}{\Ets{f(\vtheta_t^{s+1})}}
	+\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}  \nonumber\\
	&= 
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}
\end{align}
Hence:
\begin{align}
	\left|\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
	\right.&-\left.\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}\right| 
	=
	\left|\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right|  
	\nonumber\\
	&\leq
	\Tr\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right| \nonumber \\
	%
	&\leq
	\Tr\left(\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{f(\vtheta_t^{s+1})}}\right) \nonumber\\
	%
	&\leq	
	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\\
	%
	&=
	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \\
	%
	&\leq
	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2},
	\nonumber
\end{align}
where (6) is from Definition \ref{def:ideal}; (7) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (8) is from $\Ets{e_s}=0$; (9) is from Young's inequality; and (10) is from the definition of $\gradApp{\vtheta}{N}$.
\end{proof}

%Lemma 1
\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
Under Assumptions \ref{ass:bounded_score} ans \ref{ass:M2}, the expected squared norm of the true gradient $\gradJ{\vtheta_t^{s+1}}$, for appropriate choices of $\alpha_t\geq0$ and $\beta_t>0$, can be bounded as follows:
\[
	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_t\Tr\Varts{\gradApp{\tilde{\vtheta}^s}{1}}}{N\Psi_t}
	+\frac{f_t\Omega}{\Psi_t},
\]
	where
\begin{align*}
	&R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \\
	&c_{m-1} = 0, \\
	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\alpha_t^2\right) + \alpha_tL^2+\frac{\alpha_t^2L^3}{2}, \\
	&\Psi_t = \alpha_t\left(1-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right), \\
	&f_t = \alpha_t^2\frac{G^2R^2(L+2c_{t+1})}{2(1-\gamma)^2}.
\end{align*}
In particular, the following constraints on $\alpha_t$ and $\beta_t$ are sufficient:
\begin{align*}
&0\leq\alpha_t < \frac{2(1-\nicefrac{c_{t+1}}{\beta_t})}{L+2c_{t+1}} \\
&\beta_t > c_{t+1}.
\end{align*}
\end{restatable}
\begin{proof}
	We have:
	\begin{align}
	\Ets{J(\vtheta_{t+1}^{s+1})} 
	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \\
	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
	&\geq
	\Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
	&\qquad-
	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2},
\end{align}
where (11) is from the L-smoothness of $J(\vtheta)$, (12) is from the SVRPG update, and (13) is from Lemma \ref{lemma:aux0}.

@@

	+G^2\frac{R^2}{(1-\gamma)^2}\Omega\right)

@@

	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2\Omega}{2(1-\gamma)^2} \nonumber\\

@@

	-f_t\Omega,\nonumber

@@

where (17) is from (13), (18) is from (16), and (19) is from Lemma \ref{lemma:aux0}.
To complete the proof, besides rearranging terms, we have to ensure that $\Psi_t>0$ for each $t$. This gives the constraints on $\alpha_t$ and $\beta_t$.
\end{proof}
